Nail: A practical interface generator for data formats

### 2014-04-30 paper/osdi14/eval.tex
more edits
Acked-by: Julian Bangert <jbangert@acm.org>

%code size unzip 6.0
%2821 lines
\label{s:eval-effort}
To answer the second question, we implemented small example applications based on the above
grammars and compared code size with comparable applications that process data manually. We will
also compare a toy DNS server implemented in Nail with a similar toy DNS server provided as an
example for the Hammer parser generator. All code size measurements for C were performed by
SlocCount\cite{sloccount}. \XXX[code size is primitive, but we show orders of magnitude in the same language].

@@

Our DNS grammar, which is partly reproduced in Figure~\ref{fig:dns-full} consists of a 48 line  Nail
grammar and 64 lines of C implementing DNS label compression. \XXX[Make this smaller]
The grammar describes both the structure of DNS packets (36 lines) and of a simple zone-file format
(16 lines) supporting A, NS, MX and CNAME records.

@@

\paragraph{Network stack.}
\XXX[Write this up]
\subsection{Security.} 
\XXX[Add PROTOS genome CVE's. count]
 \XXX[The ZIP format has been associated with many vulnerabilities, and the PROTOS Genome project found numerous
security vulnerabilities in most implementations of ZIP and other archive formats. Furthermore,
Android security has been completely by-passed multiple times through parsing differentials in
handling ZIP files\cite{saurik-masterkey}. mention talk by Julia Wolf about zip]. 
Fuzz DNS. Fuzz ZIP. Design argument that security is hard to show.

\subsection{Performance.}
\input{fig-perf-dns}
To estimate the performance impact of using Nail grammars as opposed to hand-written grammars, we
benchmarked our DNS server against BIND \XXX[and our network stack against \texttt{lwip-tap}, a userspace
router implemented with the LWIP network stack\cite{lwip}].

\paragraph{DNS}
We compare the performance of our DNS server to the release 9.9.5 of ISC BIND 9\cite{bind8}, one of
the oldest and most popular DNS daemons, on a load that resembles the authoritative name server at a
hosting company. First, we generated domain names consisting of one or
two labels randomly selected from an English dictionary and one label that is one of three popular
Top Level Domains: ``com'', ``net'' or ``org''. Second, we randomly selected 90\% of these domains and
created a zone file that mapped these domain names to the local IP address.

Finally, we used the \texttt{queryperf} tool provided with BIND to query each domain between zero
and three times on a local instance of BIND or NailDNS. Both DNS servers were operating on a single, 
core  of an Intel i7-3610QM with 12GB of RAM on an idle system, and the benchmark tool kept at most 20
queries outstanding at once. The queryperf tool was configured to
repeat the same, randomised sequence of queries for one minute and the average throughput and
response latency were measured.  We repeated each test seven times  with 50.000 and one million
domain names, restarting each daemon in between. We also performed one initial dry run to warm the
file system cache for the zone file.
\XXX[Add 1 million results.They are largely the same]

Surprisingly, our simple Nail server outperforms the established BIND server that has seen decades
of optimisation by a factor of 3. \XXX[Why the hell?] This demonstrates that even though Nail was
not developed focusing on performance, applications built on Nail can compete with real-world
hand-written parsers.

>>>>

%code size unzip 6.0
%2821 lines
\label{s:eval-effort}
To answer the second question, we implemented small example applications based on the above
grammars and compared code size with comparable applications that process data manually. We will
also compare a toy DNS server implemented in Nail with a similar toy DNS server provided as an
example for the Hammer parser generator. All code size measurements for C were performed by
SlocCount\cite{sloccount}.

@@

Our DNS grammar, which is partly reproduced in Figure~\ref{fig:dns-full} consists of a 48 line  Nail
grammar and 64 lines of C implementing DNS label compression.
The grammar describes both the structure of DNS packets (36 lines) and of a simple zone-file format
(16 lines) supporting A, NS, MX and CNAME records.

@@

\paragraph{Network stack.}
\XXX[Write this up]
\subsection{Security.} 
We use a twofold approach to evaluate the security of applications implemented with Nail. First, we
analyse a list of CVE's related to the ZIP file format and argue how our ZIP tools based on Nail are
immune against those vulnerability classes. Second, we present the results of fuzz-testing our DNS
server.

The ZIP format has been associated with many vulnerabilities, and the PROTOS Genome project found numerous
security vulnerabilities in most implementations of ZIP and other archive formats that are directly
related to input handling.

Between May 2011 and May 2014, the CVE database\cite{cve-database} lists 83 vulnerabilities when searching for the
string ``ZIP''. Of these, 42 were completely unrelated to the archive format, for example reporting
vulnerabilities revealing Zip codes of other users. 26 vulnerabilities were related to ZIP archives,
but not to input processing, such as trusting arbitrary update ZIP files sent to a server or
vulnerabilities that can be accessed through manipulating files in a ZIP container. This includes
several vulnerabilities allowing another files to replaced while they are extracted from ZIP
files because insufficiently secure temporary directories are used. Of the 15 remaining
vulnerabilities pertaining to ZIP files, 14 vulnerabilities were related to input processing. The
 other vulnerability, which falls outside of Nails purview, concerns a ZIP packer choosing a weak
 legacy cryptosystem even when configured to use AES.

 Nail parsers are immune by design to vulnerabilities like these 14. Ten of those vulnerabilities
 are memory corruption attacks, that could also be avoided by using safer languages or better
 verification tools. Most of the buffer overflows seem to result from inconsistent offset fields
 being manipulated without checking. Because Nail's generated code checks offsets before reading and
 does not expose any untrusted pointers to the application, it protects against memory corruption
 attacks by design.

The remaining four vulnerabilities do not involve memory corruption, but rather parser ambiguities.
Two of them relate to implementations mis-parsing ZIP files, particularly in virus scanners. One
other vulnerability is in a Python ZIP library, which uses both the size field in the file entry and
in the header, without checking for their consistency. This results in a denial-of-service through
endless loops or incorrectly extracted ZIP files.

A similar attack was used by the fourth vulnerability, the infamous Android master key bug
\cite{saurik-masterkey}, that completely by-passed Android security multiple times through parser
differentials between the ZIP handler that checks signatures for privileged applications and the ZIP
implementations that ultimately extracts those files. Thus, valid  application bundles could be
modified to include malicious applications without breaking their signatures. Instead of both
components containing an ad-hoc archive implementation, a single Nail grammar would be reusable
across both implementations. 
\subsection{Performance.}
\input{fig-perf-dns}
To estimate the performance impact of using Nail grammars as opposed to hand-written grammars, we
benchmarked our DNS server against BIND.

We compare the performance of our DNS server to the release 9.9.5 of ISC BIND 9\cite{bind8}, one of
the oldest and most popular DNS daemons, on a load that resembles the authoritative name server at a
hosting company. First, we generated domain names consisting of one or
two labels randomly selected from an English dictionary and one label that is one of three popular
Top Level Domains: ``com'', ``net'' or ``org''. Second, we randomly selected 90\% of these domains and
created a zone file that mapped these domain names to the local IP address.

Finally, we used the \texttt{queryperf} tool provided with BIND to query each domain between zero
and three times on a local instance of BIND or NailDNS. Both DNS servers were operating on a single, 
core  of an Intel i7-3610QM with 12GB of RAM on an idle system, and the benchmark tool kept at most 20
queries outstanding at once. The queryperf tool was configured to
repeat the same, randomised sequence of queries for one minute and the average throughput and
response latency were measured.  We repeated each test seven times  with 50.000
domain names, restarting each daemon in between. We also performed one initial dry run to warm the
file system cache for the zone file. We repeated the tests with 1 million domain names and found
that the servers performed almost identically.

Surprisingly, our simple Nail server outperforms the established BIND server that has seen decades
of optimisation by a factor of 3.  This demonstrates that even though Nail was
not developed focusing on performance, applications built on Nail can compete with real-world
hand-written parsers.

### 2014-04-30 paper/osdi14/impl.tex
more edits
Acked-by: Julian Bangert <jbangert@acm.org>

The current prototype of the Nail parser generator supports the C programming
language and top-down parsers. Options for C++ STL data models and emitting
Packrat parsers~\cite{packrat-parsing:icfp02} are under development. In
this section, we will discuss some particular features of our parser
implementation.

@@

\paragraph{Bootstrapping.}

To demonstrate the feasibility of the Nail parser generator, our parser
generator uses a Nail parser to recognize Nail grammars. A superset of the
grammar language described in this paper is implemented in 100 lines of Nail,
which feed into about 1,000 lines of C++ that implement the actual parser
generator. Bootstrapping is supported via an implementation of the Nail language
in the Lemon parser generator, a variant of Yacc.

>>>>

The current prototype of the Nail parser generator supports the C programming
language and top-down parsers. The implementation parses Nail grammars with Nail itself, using a 130
line Nail grammar feeding into a 2000 line C++ program that implements the parser. Bootstrapping is
performed with a subset of the grammar implemented using conventional grammars. 
Options for C++ STL data models and emitting
Packrat parsers~\cite{packrat-parsing:icfp02} are under development. In
this section, we will discuss some particular features of our parser
implementation.

@@

### 2014-04-30 paper/osdi14/relwk.tex
more edits
Acked-by: Julian Bangert <jbangert@acm.org>

@@

\paragraph{Language safety.}

The second class is logic errors, where application code misinterprets
input data.  This can lead to serious security consequences when two
systems disagree on the meaning of a network packet or a signed message,
as in iOS\footnote{The XNU kernel and the user code-signing verifier
interpret executable metadata differently, so the code signature
sees different bytes at a virtual address than the executable that
runs.}~\cite{evaders6} and Android\footnote{Android applications
are distributed as .zip files. Signatures are verified with a Java
program, but the program is extracted with a C program.  The Java
program interprets all fields as signed, whereas the C program
treats them as unsigned, allowing one to replace files in a signed
archive, thereby undermining Android's security model.}~\cite{saurik-masterkey}\XXX{Duplicate, we
already mention this in the eval} code signing and even the X.509 protocol
underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}.  These mistakes are
highly application-specific, and are difficult to mitigate using existing
techniques, and these mistakes can occur even in high-level languages
that guarantee memory safety.
By allowing developers to specify their data format just once, Nail
avoids logic errors and inconsistencies in parsing and output generation.

A subclass of logic errors are so-called \textit{weird machines}, where
implementation side effects or under-specified parser behavior leads to a
protocol or data format inadvertently becoming a Turing-complete execution
environment, even though the original grammar did not require it.
Frequently, this execution environment can either
then directly manipulate data in unwanted ways or be used to make
exploiting another bug feasible.\footnote{For example, by compiling a
return-oriented-programming exploit from code fragments discovered on
the fly.} Examples include x86 page tables~\cite{bangert2013page}, and
ELF symbols and relocations~\cite{shapiro2013weird}. In the offensive
research community, this has been generalized into treating a program
as a \textit{weird machine}~\cite{bratus2011exploit} that operates on
an input, analogous to a virtual machine operating on bytecode.
Nail avoids these problems by having the parser precisely match the
specified grammar, eliminating under-specified behavior.

\paragraph{Parsing frameworks.}
Proper input recognition has  been shown to be an excellent way of
eliminating malicious inputs. In one case, a  PDF parser implemented in
Coq could eliminate over $95\%$ of known malicious PDFs~\cite{Bogk-PDF}.
However, manually writing parser code does not scale to the number of
file formats  and protocols in existence and might result in parser code
tied to one specific application.

@@

Nail improves upon Hammer in three ways. First, Nail  generates output besides recognizing input.
Second, Nail does not require the programmer to write potentially insecure semantic actions. Last,
Nail's structural dependencies allow it to work with protocols Hammer cannot recognize, such as
protocols with offset fields or length fields (Hammer has extremely limited support for length
fields: it can parse arrays immediately preceded by their length).

>>>>

\paragraph{Motivation.}
\XXX[Move to the intro, into its own section, or kill?]

The second class is logic errors, where application code misinterprets
input data.  This can lead to serious security consequences when two
systems disagree on the meaning of a network packet or a signed message,
as in iOS\footnote{The XNU kernel and the user code-signing verifier
interpret executable metadata differently, so the code signature
sees different bytes at a virtual address than the executable that
runs.}~\cite{evaders6} and Android~\cite{saurik-masterkey}code signing and even the X.509 protocol
underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}.  These mistakes are
highly application-specific, and are difficult to mitigate using existing
techniques, and these mistakes can occur even in high-level languages
that guarantee memory safety.
By allowing developers to specify their data format just once, Nail
avoids logic errors and inconsistencies in parsing and output generation.

% A subclass of logic errors are so-called \textit{weird machines}, where
% implementation side effects or under-specified parser behavior leads to a
% protocol or data format inadvertently becoming a Turing-complete execution
% environment, even though the original grammar did not require it.
% Frequently, this execution environment can either
% then directly manipulate data in unwanted ways or be used to make
% exploiting another bug feasible.\footnote{For example, by compiling a
% return-oriented-programming exploit from code fragments discovered on
% the fly.} Examples include x86 page tables~\cite{bangert2013page}, and
% ELF symbols and relocations~\cite{shapiro2013weird}. In the offensive
% research community, this has been generalized into treating a program
% as a \textit{weird machine}~\cite{bratus2011exploit} that operates on
% an input, analogous to a virtual machine operating on bytecode.
% Nail avoids these problems by having the parser precisely match the
% specified grammar, eliminating under-specified behavior.

@@

\XXX[Parsifal at langsec '14]
\paragraph{Related Work}
Proper input recognition has  been shown to be an excellent way of
eliminating malicious inputs. In one case, a  PDF parser implemented in
Coq could eliminate over $95\%$ of known malicious PDFs~\cite{Bogk-PDF}.
However, manually writing parser code does not scale to the number of
file formats  and protocols in existence and might result in parser code
tied to one specific application.

@@

Nail improves upon Hammer in three ways. First, Nail  generates output besides recognizing input.
Second, Nail does not require the programmer to write potentially insecure semantic actions. Last,
Nail's structural dependencies and transformations allow it to work with protocols Hammer cannot recognize, such as
protocols with offset fields or length fields (Hammer has extremely limited support for length
fields: it can parse arrays immediately preceded by their length).
